---
title: "100-basics"
output: html_document
date: "2023-07-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
```

```{r}
meta<-read.csv("meta/PUTSampleBarcodes_062623.csv") %>% separate(SampleName, into=c("Trib","Species","Method","Year","Number"), remove = FALSE) 
meta
```
```{r}
meta %>% group_by(Year, Method) %>% summarize(Count=n())
```

Lots of samples over several years!         


Have raw data.  Looks like nine plates sequenced on two lanes. 
230720_HiSeq_Run751      
HS: PE 150+8+8+150      

Lane 1 Project_ASEF_BMAG077_L1         
PUT014RAD	CAAAAG
PUT015RAD	CAACTA
PUT016RAD	CAGGCG
PUT017RAD	CATGGC
PUT018RAD	CATTTT
PUT019RAD	CCAACA
PUT020RAD	ACAGTG
PUT021RAD	GCCAAT
PUT022RAD	TAATCG

Lane 2     Project_ASEF_BMAG077_L2


PUT014RAD	CAAAAG
PUT015RAD	CAACTA
PUT016RAD	CAGGCG
PUT017RAD	CATGGC
PUT018RAD	CATTTT
PUT019RAD	CCAACA
PUT020RAD	ACAGTG
PUT021RAD	GCCAAT
PUT022RAD	TAATCG

Putting in data/raw     

check md5, combine lanes.       
See 100.1-combine.sh     
module load parallel     
srun -p high -t 02:00:00 --nodes=1 --ntasks-per-node=1 --cpus-per-task=9 parallel -j 9 < 100.1-combine.sh    

Set up some scripts to demultiplex in scripts    

