---
title: "101-first-call"
output: html_document
date: "2023-08-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(RcppCNPy)
library(viridis)
```

```{r}
meta<-read_csv("meta/putah-creek-meta-with-counts.csv") %>% mutate(Path=paste0("data/split/",SampleName,".sort.flt.bam"))
meta %>% group_by(Method, Year) %>% summarise(Count=n())
```


Let's get a subset of fish and try some things. I suspect the CS samples to be problematic.    

```{r}
df1<-meta %>% filter(Filtered > 1e5) 
df1 %>% group_by(Method, Year)  %>% mutate(Mean=mean(Filtered), Median=median(Filtered), Max=max(Filtered)) %>% select(Method, Year, Mean, Median, Max) %>%
  mutate(Count=n()) %>% unique() %>% arrange(Method,Year)
```

Can try df1 here....

```{r}
write_tsv(df1 %>% select(Path), col_names = FALSE, file="bamlists/243.bamlist")
write_tsv(df1 %>% select(SampleName), col_names = FALSE, file="bamlists/243-sample-names.txt")

```

Generate covariance matrix with -doIBS

```{sh, eval=FALSE}
srun -p high -t 3:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 16  \
-bam bamlists/243.bamlist -r NC_056429.1 \
-ref  ~/genomes/chinook-2/GCF_018296145.1_Otsh_v2.0_genomic.fna \
-minInd 122 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/101/243-ibs > outputs/101/243-ibs.out 2> outputs/101/243-ibs.err &
```

with minInd 183
	-> Number of sites retained after filtering: 0 

with minInd 122
	-> Number of sites retained after filtering: 134 

```{sh, eval=FALSE}
srun -p high -t 3:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 16  \
-bam bamlists/243.bamlist  -rf meta/chrom-list.txt \
-ref  ~/genomes/chinook-2/GCF_018296145.1_Otsh_v2.0_genomic.fna \
-minInd 122 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/101/243-ibs > outputs/101/243-ibs.out 2> outputs/101/243-ibs.err &

srun -p high -t 3:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 16  \
-bam bamlists/243.bamlist  -rf meta/chrom-list.txt \
-ref  ~/genomes/chinook-2/GCF_018296145.1_Otsh_v2.0_genomic.fna \
-minInd 182 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/101/243-ibs-75 > outputs/101/243-ibs-75.out 2> outputs/101/243-ibs-75.err &

srun -p high -t 3:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 16  \
-bam bamlists/243.bamlist  -rf meta/chrom-list.txt \
-ref  ~/genomes/chinook-2/GCF_018296145.1_Otsh_v2.0_genomic.fna \
-minInd 219 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/101/243-ibs-90 > outputs/101/243-ibs-90.out 2> outputs/101/243-ibs-90.err &

## GLS
srun -p high -t 3:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 16  \
-bam bamlists/243.bamlist -rf meta/chrom-list.txt \
-ref ~/genomes/chinook-2/GCF_018296145.1_Otsh_v2.0_genomic.fna \
-minInd 182 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-minMaf 0.05 \
-out outputs/101/243-gl-75 > outputs/101/243-gl-75.out 2> outputs/101/243-gl-75.err &
```

50% doIBS      
	-> Number of sites retained after filtering: 3443     
75% doIBS      
	-> Number of sites retained after filtering: 552      
	
75% beagle
  -> Number of sites retained after filtering: 552      

```{r}
pcm<-df1
m <- as.matrix(read.table("outputs/101/243-ibs-75.covMat"))
eig <- eigen(m)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)


covs<-eig$vectors[,1:3] %>% as_tibble() %>% bind_cols(pcm) 

#text12<-covs %>% select(Run, Pop, V1, V2) %>%
 # group_by(Pop) %>% summarize(Count=n(), x=mean(V1), y=mean(V2))
covs$Year<-as.factor(covs$Year)
ggplot(covs) +
  geom_point(aes(x=V1, y=V2, fill=Year, shape=Method), alpha=0.75) +
#  geom_text_repel(data=text12, aes(x=x, y=y, label=Pop), max.overlaps = Inf) +
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = "")) +
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = "")) +
  theme_bw() +
  theme(panel.grid=element_blank()) +
  scale_fill_viridis_d(option="H") +
  guides(fill = guide_legend(override.aes = list(pch=22))) +
  scale_shape_manual(values=c(21,23))


```


```{r}
ggplot(covs) +
  geom_point(aes(x=V1, y=V3, fill=Year, shape=Method), alpha=0.75) +
#  geom_text_repel(data=text12, aes(x=x, y=y, label=Pop), max.overlaps = Inf) +
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = "")) +
  ylab(paste("PC3", " ", round((100*var[3]),2), "%", sep = "")) +
  theme_bw() +
  theme(panel.grid=element_blank()) +
  scale_fill_viridis_d(option="H") +
  guides(fill = guide_legend(override.aes = list(pch=22))) +
  scale_shape_manual(values=c(21,23))
```
Seems fine.    

What about relationships? can use older PCAngsd or ngsRelate...

PCAngsd goes something like this (using 3,443 sites and 243 samples)

```{sh, eval=FALSE}
conda activate py2

python $HOME/pcangsd/pcangsd.py -beagle outputs/101/243-ibs.beagle.gz -kinship -o outputs/101/243-ibs -threads 10

python $HOME/pcangsd/pcangsd.py -beagle outputs/101/243-ibs.beagle.gz -relate outputs/101/243-ibs.kinship.npy -o outputs/101/243-ibs-kin-removed -threads 10


srun -p high -t 1:00:00 --nodes=1 python $HOME/pcangsd/pcangsd.py -beagle outputs/101/243-ibs.beagle.gz -kinship -o outputs/101/243-ibs-10 -threads 10 -minMaf 0.10

srun -p high -t 1:00:00 --nodes=1 python $HOME/pcangsd/pcangsd.py -beagle outputs/101/243-ibs-75.beagle.gz -kinship -o outputs/101/243-ibs-75 -threads 10 
srun -p high -t 1:00:00 --nodes=1 python $HOME/pcangsd/pcangsd.py -beagle outputs/101/243-ibs-90.beagle.gz -kinship -o outputs/101/243-ibs-90 -threads 10 

```

Keeping 243 individuals after filtering (removing 0)

ruh-roh

Higher MAF
Number of sites after MAF filtering (0.1): 2413

     
```{r}
kin<-npyLoad("outputs/101/243-ibs-75.kinship.npy")

#Convert upper half of matrix to data frame compatible 
kinrind<-which(upper.tri(kin, diag = TRUE), arr.ind = TRUE)
kin<-cbind(kinrind, kin[kinrind])

kindf<-as_tibble(kin) %>%
  rename(Ind1=row, Ind2=col, Kinship=V3) %>%
  filter(Ind1 != Ind2) %>%
  mutate(Mean=mean(Kinship))


#Can I replace Ind1 and Ind2 with individual labels???
df<-kindf %>% left_join(pcm %>% mutate(Individual=1:n()), by=c("Ind1"="Individual")) %>% dplyr::rename(Year1=Year, Method1=Method, SampleName1=SampleName) %>%
  left_join(pcm %>% mutate(Individual=1:n()), by=c("Ind2"="Individual")) %>% dplyr::rename(Year2=Year, Method2=Method,  SampleName2=SampleName)

kinship<-df %>% select(Ind1, Ind2, Kinship, SampleName1, SampleName2, Method1, Year1, Method2, Year2)
```

```{r}
#Ahh, and what are our "related" individuals
kinship %>% filter(Kinship>0.0625)

```

high<-df %>% filter(Kinship>0.0625)

#filter to the same pops and get those locations with closely related individuals
popdf<-df %>% filter(Location1==Location2) %>% mutate(Population=Location1) %>%
  filter(Population %in% high$Location1)

```{r}
#Plotting 
cols<-viridis(6)


ggplot(kinship)+geom_histogram(aes(x=Kinship), bins=50, fill="grey", color="grey") +
  geom_vline(xintercept =.25, col=cols[1]) + #Parent offspring/full sibs
  geom_vline(xintercept =.125, col=cols[2]) + #2nd degree
  geom_vline(xintercept =.0625, col=cols[3]) + # 3rd
  geom_vline(xintercept = .0313, col=cols[4]) +
  geom_vline(xintercept = .0156, col=cols[5]) +
  geom_vline(xintercept = 0.0, col=cols[6]) +
  scale_x_continuous(breaks=c(0,0.0156, 0.0313, 0.0625, 0.125, 0.25),
                     labels=c("0 / Unrelated","5th Deg.","4th Deg.","3rd Deg.", "2nd Deg.", "1st Deg.")) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust=1, angle=45)) +
  theme(panel.grid=element_blank()) +
  ylab("Count") +
  xlab("Kinship Coefficient")
```
#  facet_wrap(Population ~ ., ncol=3, scales="free_y")


#Looking at more fish



```{r}
df2<-meta %>% filter(Filtered > 5e4) 
df2 %>% group_by(Method, Year)  %>% mutate(Mean=mean(Filtered), Median=median(Filtered), Max=max(Filtered)) %>% select(Method, Year, Mean, Median, Max) %>%
  mutate(Count=n()) %>% unique() %>% arrange(Method,Year)
```



```{r}
write_tsv(df2 %>% select(Path), col_names = FALSE, file="bamlists/383.bamlist")
```


```{sh, eval=FALSE}
srun -p high -t 3:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 16  \
-bam bamlists/383.bamlist  -rf meta/chrom-list.txt \
-ref  ~/genomes/chinook-2/GCF_018296145.1_Otsh_v2.0_genomic.fna \
-minInd 192 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/101/383-ibs > outputs/101/383-ibs.out 2> outputs/101/383-ibs.err &

srun -p high -t 3:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 16  \
-bam bamlists/383.bamlist  -rf meta/chrom-list.txt \
-ref  ~/genomes/chinook-2/GCF_018296145.1_Otsh_v2.0_genomic.fna \
-minInd 287 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/101/383-ibs-75 > outputs/101/383-ibs-75.out 2> outputs/101/383-ibs-75.err &

srun -p high -t 3:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 16  \
-bam bamlists/383.bamlist  -rf meta/chrom-list.txt \
-ref  ~/genomes/chinook-2/GCF_018296145.1_Otsh_v2.0_genomic.fna \
-minInd 345 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/101/383-ibs-90 > outputs/101/383-ibs-90.out 2> outputs/101/383-ibs-90.err &

## GLS
srun -p high -t 3:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 16  \
-bam bamlists/383.bamlist -rf meta/chrom-list.txt \
-ref ~/genomes/chinook-2/GCF_018296145.1_Otsh_v2.0_genomic.fna \
-minInd 287 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-minMaf 0.05 \
-out outputs/101/383-gl-75 > outputs/101/383-gl-75.out 2> outputs/101/383-gl-75.err &


##



srun -p high -t 1:00:00 --nodes=1 python $HOME/pcangsd/pcangsd.py -beagle outputs/101/383-ibs.beagle.gz -kinship -o outputs/101/383-ibs -threads 10 &
srun -p high -t 1:00:00 --nodes=1 python $HOME/pcangsd/pcangsd.py -beagle outputs/101/383-ibs-75.beagle.gz -kinship -o outputs/101/383-ibs-75 -threads 10 &
srun -p high -t 1:00:00 --nodes=1 python $HOME/pcangsd/pcangsd.py -beagle outputs/101/383-ibs-90.beagle.gz -kinship -o outputs/101/383-ibs-90 -threads 10 &
srun -p high -t 1:00:00 --nodes=1 python $HOME/pcangsd/pcangsd.py -beagle outputs/101/383-gl-75.beagle.gz -kinship -o outputs/101/383-gl-75 -threads 10 &

```
	-> Number of sites retained after filtering: 1359          
  -> Number of sites retained after filtering: 335     
	-> Number of sites retained after filtering: 160          


Eyeball for sanity.     

```{r}
pcm<-df2
m <- as.matrix(read.table("outputs/101/383-ibs-75.covMat"))
eig <- eigen(m)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)


covs<-eig$vectors[,1:3] %>% as_tibble() %>% bind_cols(pcm) 

#text12<-covs %>% select(Run, Pop, V1, V2) %>%
 # group_by(Pop) %>% summarize(Count=n(), x=mean(V1), y=mean(V2))
covs$Year<-as.factor(covs$Year)
ggplot(covs) +
  geom_point(aes(x=V1, y=V2, fill=Year, shape=Method), alpha=0.75) +
#  geom_text_repel(data=text12, aes(x=x, y=y, label=Pop), max.overlaps = Inf) +
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = "")) +
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = "")) +
  theme_bw() +
  theme(panel.grid=element_blank()) +
  scale_fill_viridis_d(option="H") +
  guides(fill = guide_legend(override.aes = list(pch=22))) +
  scale_shape_manual(values=c(21,23))


```



```{r}
kin<-npyLoad("outputs/101/383-ibs-75.kinship.npy")

#Convert upper half of matrix to data frame compatible 
kinrind<-which(upper.tri(kin, diag = TRUE), arr.ind = TRUE)
kin<-cbind(kinrind, kin[kinrind])

kindf<-as_tibble(kin) %>%
  rename(Ind1=row, Ind2=col, Kinship=V3) %>%
  filter(Ind1 != Ind2) %>%
  mutate(Mean=mean(Kinship))


#Can I replace Ind1 and Ind2 with individual labels???
df<-kindf %>% left_join(pcm %>% mutate(Individual=1:n()), by=c("Ind1"="Individual")) %>% dplyr::rename(Year1=Year, Method1=Method, SampleName1=SampleName) %>%
  left_join(pcm %>% mutate(Individual=1:n()), by=c("Ind2"="Individual")) %>% dplyr::rename(Year2=Year, Method2=Method,  SampleName2=SampleName)

kinship<-df %>% select(Ind1, Ind2, Kinship, SampleName1, SampleName2, Method1, Year1, Method2, Year2)
```

```{r}
#Ahh, and what are our "related" individuals
kinship %>% filter(Kinship>0.0625) %>% arrange(-Kinship)

```

high<-df %>% filter(Kinship>0.0625)

#filter to the same pops and get those locations with closely related individuals
popdf<-df %>% filter(Location1==Location2) %>% mutate(Population=Location1) %>%
  filter(Population %in% high$Location1)

```{r}
#Plotting 
cols<-viridis(6)


ggplot(kinship)+geom_histogram(aes(x=Kinship), bins=50, fill="grey", color="grey") +
  geom_vline(xintercept =.25, col=cols[1]) + #Parent offspring/full sibs
  geom_vline(xintercept =.125, col=cols[2]) + #2nd degree
  geom_vline(xintercept =.0625, col=cols[3]) + # 3rd
  geom_vline(xintercept = .0313, col=cols[4]) +
  geom_vline(xintercept = .0156, col=cols[5]) +
  geom_vline(xintercept = 0.0, col=cols[6]) +
  scale_x_continuous(breaks=c(0,0.0156, 0.0313, 0.0625, 0.125, 0.25),
                     labels=c("0 / Unrelated","5th Deg.","4th Deg.","3rd Deg.", "2nd Deg.", "1st Deg.")) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust=1, angle=45)) +
  theme(panel.grid=element_blank()) +
  ylab("Count") +
  xlab("Kinship Coefficient")
```


Hmm... Are these reasonable?    
